\documentclass{article}
\usepackage{Self_Style}

\title{UChicago REU Notes Week 2}
\author{Zih-Yu Hsieh}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\section{Week 2}

\subsection{Intro}
For this portion, I want to start with some basic definitions covered by professor May in his office, about the definition of spectra / prespectra / generalized cohomology theory. I'll probably skip some of the details in the intro to make sure I can keep on track. (Also, I probably involve quite a lot of errors in between...I'll try to fix them)
\subsubsection{Brown Representability Theorem, and Generalized Cohomology}
Let $h\mathcal{C}$ denotes the homotopy category of based spaces (that admits a CW structure, as on them weak equivalences and homotopy equivalence are the same), then a functor $F:h\mathcal{C}^\Op\rightarrow \Ab$ is \emph{representable}, if there exists a based space $Y\in h\mathcal{C}$, such that $F\cong [\_,Y]$.
\begin{exmp}
    The \emph{reduced cohomology functor} $\tilde{H}^n(\_;A)$ is one, for all $n\geq 0$ and a fixed abelian group $A$ (as coefficient). Let $K(A,n)$ be the \emph{Eilenberg-Maclane Space} of abelian group $A$ and index $n\geq 0$ (i.e. the $n$th homotopy group $\pi_n (K(A,n))=A$), then the following group isomorphism holds:
    \[\tilde{H}^n(X;A) \cong [X, K(A,n)]\]
    Using more adjunction property, since $\pi_n (A,n) = A = \pi_{n+1}(K(A,n+1)) = [S^{n+1}, K(A,n+1)] = [S^n \wedge S^1, K(A,n+1)] \cong [S^n, \Omega K(A,n+1)]= \pi_n (\Omega K(A,n+1))$, there is a homotopy equivalence $K(A,n)\xrightarrow{\simeq} \Omega K(A,n+1)$.
\end{exmp} 

(Yeah, need to understand more about the details)

So, the regular cohomology we see on spaces, are one type of representable functors.

The classification of all such functors is given by \emph{Brown Representability Theorem}:
\begin{thm}
Suppose the functor $F:h\mathcal{C}^\Op\rightarrow \Sets$ satisfies:
\begin{itemize}
    \item $F(\bigvee_\alpha X_\alpha)\cong \prod_\alpha F(X_\alpha)$ (i.e. maps coproduct to product).
    \item $F$ maps \emph{homotopy pushouts} (I believe this is the reduced mapping cylinder?) to \emph{weak pullbacks} (omit the uniqueness of maps). This condition is equivalent to \emph{Mayer-Vietoris axiom} (which seems like a weaker form of the gluing axiom in sheaf theory, and restrict to CW subcomplexes instead of open sets).
\end{itemize}
Then, $F$ is representable by some $CW$ complex $C$, or $F\cong [\_,C]$.
\end{thm}
The converse for the theorem also holds, as $[\_,C]$ also satisfies the above axioms (need more details here). 

\hfil

Here is a definition of generalized cohomology:

\begin{defn}
A \emph{generalized cohomology}, is a sequence of spaces $\{E_n\}_{n\geq 0}$, with homotopy equivalences $\{\epsilon_n:E_n \xrightarrow{\simeq} \Omega E_{n+1}\}_{n\geq 0}$. 

In a more explicit form, $E^n(X) := [X,E_n]$ has all $n\geq 0$ satisfies $E^n(X) = [X,E_n]\cong [X,\Omega E_{n+1}]$, which is naturally a group (and an abelian group for $n\geq 2$, due to the abelian group structure on $[X,\Omega^m Y]$ for $m\geq 2$).
\end{defn}
I remembered the motivation is that, by Brown Representability Theorem, if a sequence of representable functors $E^*:h\mathcal{C}^\Op\rightarrow \Sets$ satisfy similar properties to ordinary cohomology (for instance, the connecting morphism $E^n\rightarrow E^{n+1}$, and excision axiom?), then Brown Representability Theorem guarantees it must be the above form. So, they're precisely all the generalized cohomology.

\hfil

\subsubsection{$\Omega$-Prespectra, Prespectra, and Spectra}
When talking about generalized cohomology, we have a sequence of based spaces $\{E_n\}_{n\geq 0}$, together with homotopy equivalences $\epsilon_n:E_n \xrightarrow{\simeq}\Omega E_{n+1}$, which is a prototype of what's called \emph{$\Omega$-Prespectra}.

In general, fixing a based space $X\in h\mathcal{C}$, the sequence $\{\Sigma^n X\}_{n\geq 0}$ has a collection of natural maps $\{\id_n:\Sigma (\Sigma^n X)\rightarrow \Sigma^{n+1}X\}_{n\geq 0}$, which the adjunction $(\Sigma, \Omega)$ guarantees a collection of maps $\{\eta_n:\Sigma^n X\rightarrow \Omega (\Sigma^{n+1}X)\}_{n\geq 0}$. In general these maps are not homotopy equivalence, so these type of sequences are a type of \emph{Prespectra}, called \emph{suspension prespectra}. More generally, \emph{Prespectra} is a sequence of spaces $\{X_n\}_{n\geq 0}$, with morphisms $\{\mu_n:X_n\rightarrow \Omega X_{n+1}\}_{n\geq 0}$.

For the strongest form possible, if such map $\epsilon_n:E_n\rightarrow \Omega E_{n+1}$ is a homeomorphism, then these are called \emph{Spectra}.

\hfil

Professor May restricts to three specific categories: 
\begin{itemize}
    \item Let $\mathcal{P}$ denotes the \emph{category of prespectra}.
    \item Let $\mathcal{Q}$ denotes the \emph{category of inclusion prespectra}, meaningeach defining morphism $\mu_n:X_n\rightarrow \Omega X_{n+1}$ is an inclusion.
    \item Let $\mathcal{S}$ denotes the \emph{category of spectra}.
\end{itemize}
For each category, we require the morphism $\phi:X_\cdot\rightarrow T_\cdot$ to be a sequence of based maps $\phi_n:X_n\rightarrow T_n$, such that the following diagram commutes:
\[\xymatrix{
    \Sigma X_n \ar[d]_-{\mu'_{X_n}} \ar[r]^-{\Sigma\phi_n} & \Sigma T_n \ar[d]^-{\mu'_{T_n}}\\
    X_{n+1} \ar[r]_-{\phi_{n+1}} & T_{n+1}
}\]
where the map $\Sigma X_n\rightarrow X_{n+1}$ is based on the adjoint property $[\Sigma X_n,X_{n+1}]\cong [X_n,\Omega X_{n+1}]$. Here professor May had talked about how the prototype of the morphism only requires it to commute up to homotopy, but turns out it's too weak as a condition.

\hfil

With these morphisms, there is an obvious forgetful functor $\mathcal{S}\rightarrow \mathcal{Q}\rightarrow \mathcal{P}$. It seems to satisfy all the conditions for Freyd's Adjoint Functor Theorem to hold, both of the forgetful functors have a left adjoint.

Here, the left adjoint of $\mathcal{Q} \rightarrow \mathcal{P}$ is hard to construct, yet there is a clear construction for the left adjoint of $\mathcal{S}\rightarrow \mathcal{Q}$, denoted as $L:\mathcal{Q}\rightarrow\mathcal{S}$, by $T_\cdot \mapsto (LT)_\cdot$, where $(LT)_n := \colim_{q}\Omega^{n+q}T_q$ (due to the inclusion $T_q\rightarrow \Omega T_{q+1}$, it induces $\Omega^{n+q}T_q\rightarrow \Omega^{n+(q+1)}T_{q+1}$, hence having a sequence of direct system).

Professor May had claimed the homeoomorphism $(LT)_n \cong \Omega (LT)_{n+1}$ due to the fact that $\Omega$ commutes with colimit (I do need to work out the detail, as normally $\Omega$ as a right adjoint of $\Sigma$ would be thought of preserving limits instead of colimits).

\newpage

Most part of the following notesare from professor May's \emph{Infinite Loop Space Theory} paper 

(https://www.math.uchicago.edu/~may/PAPERS/18.pdf).

\subsection{Spectra, Infinite Loop Spaces, and Stabilization Functor}
If require spectra to be a sequence of spaces $E=\{E_n\}_{n\geq 0}$ with homeomorphism $E_n\cong \Omega E_{n+1}$, we have the following definition:
\begin{defn}
    A space $X$ that has a homotopy type of $E_0$ for some spectra $E$ is called an \emph{infinite loop space}.

    If $X,Y$ are infinite loop spaces of spectra $E,F$ respectively, then a map $f:X\rightarrow Y$ is caled an \emph{infinite loop map}, if it's induced by some morphism of spectra $\tilde{f}:E\rightarrow F$.
\end{defn}

Also, we have the notion of stable homotopy groups:
\begin{defn}
    Given a space $X$, its\emph{$n$th stable homotopy group} $\pi_n^s (X) := \colim_j \pi_{n+j}(\Sigma^j X)$. 
    
    Here, the colimit is due to the natural map $\Sigma(\Sigma^{j}X)\rightarrow \Sigma^{j+1}X$, inducing $\Sigma^j X\rightarrow \Omega (\Sigma^{j+1}X)$. Hence, the following isomorphism holds:
    \[\pi_k (\Sigma^j X)\rightarrow \pi_k (\Omega(\Sigma^{j+1}X)) =[\Sigma^k S^0, \Omega (\Sigma^{j+1}X)]\cong [\Sigma^{k+1}S^0, \Sigma^{j+1}X] = \pi_{k+1}(\Sigma^{j+1}X)\]
    Hence, fixing $k:= n+j$, this induces a map $\pi_{n+j}(\Sigma^j X)\rightarrow \pi_{n+(j+1)}(\Sigma^{j+1}X)$.
\end{defn}
I remembered part of the reason this is called stable homotopy group, is because for large enough $j$ this direct system in fact stabilizes.

\hfil

For stabilization, it's a functor $Q X := \colim_q \Omega^q \Sigma^q X$. Which, it has the following isomorphism on homotopy groups:
\[\pi_n (Q X)\cong \pi^s_n (X)\]

\end{document}